Integer-N PFD/CP-based PLL - System-Level simulation

### 2025-05-07 Examples/Example3_4_SimulationandEstimationofLoopSettlingTimes/article/article.qmd
Finishing report of example 3.4.

f0=2.4e9
f=2.4835e9
t_s = 225e-6
f_ref = 1e6
```

# Introdução
The intention in this report is the designing of an integer-N synthesizer that operates at `{python} format(f0)`Hz to `{python} format(f)`Hz. The synthesizer must to settle for a frequency step in a settling time of, $t_s=$ `{python} format(t_s)`s. The channel spacing is `{python} format(f_ref)`Hz and as it's using an integer-N synthesizer, the channel space is equal to the frequency reference.

Using a equation discussed in the previous report, it's possible to get the natural frequency, $\omega_n$,

$$
t_s\approx\frac{-\ln\left(0.01\sqrt{1-\zeta^2}\right)}{\zeta}\frac{1}{\omega_n}
$$

zeta=0.707
omega_n__t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta)*(-1)
omega_n=omega_n__t_s/t_s
f_n=omega_n/(2*math.pi)
```

Using an arbitrary value for $\zeta=$ `{python} zeta`. With that constant, the natural frequency would be $\omega_n=$ `{python} format_with_pi_latex(omega_n)`rad/s, or $f_n=$ `{python} format(f_n)`Hz.

Starting from $omega_n$ and $zeta$, it's possible to get the 3-dB bandwidth, $\omega_{3\text{dB}}$ as follows,

$$
\omega_{3\text{dB}}=\omega_n\sqrt{1+2\zeta^2+\sqrt{4\zeta^4+4\zeta^2+2}}
$$

omega_bw = omega_n * math.sqrt(1+2*zeta**2+math.sqrt(4*zeta**4+4*zeta**2+2))
f_bw = omega_bw/(2*math.pi)
```

Using that relation, it results in $f_{3\text{dB}}=$ `{python} format(f_bw)`Hz.

How the relation between $f_\text{ref}\gg f_{3\text{dB}}$, thus the system isn't affected for the discretization process (sampling). So the continuous analysis is useful for this example.

```{python}
#| echo: false
N0=f0/f_ref
N=f/f_ref
```

As it's an integer-N and the reference frequency $f_\text{ref}=$ `{python} format(f_ref)`Hz, the so $N$ should be from `{python} round(N0, 0)` to `{python} round(N, 0)`.

@@

f_range=f-f0
tuning_range_percent = 9.6 
tuning_range = f0/tuning_range_percent
V=2
Kvco = 250e6 # Hz/V
Kvco_ang = Kvco*2*math.pi
```

To the VCO, it's needed to cover a tuning range of  `{python} format(f_range)`Hz.

Due to process variations, it is necessary to ensure that the VCO has an additional tuning margin. Therefore, a total tuning range of approximately `{python} round(tuning_range_percent, 0)`%, of the center frequency, around `{python} format(tuning_range)`Hz, is recommended. And assuming a voltage supply of `{python} format(V)`V, a VCO constant of `{python} format(Kvco)`Hz/V `{python} format(Kvco_ang)`rad/s/V is expected. So the free running frequency of the VCO should be in minimum frequency. The middle of the frequencies, $f_c$,

$$
f_c=\frac{f+f_0}{2}
$$

f_c=(f+f0)/2
fvco=f_min=f_c-tuning_range
f_max=f_c+tuning_range

f_example_1=2.401e9
```

So it results in $f_c=$ `{python} format(f_c)`Hz. The minimum frequency is $f_\text{min}=$ `{python} format(f_min)`Hz and the maximum is `{python} format(f_max)`Hz. The free running frequency will be equals to $f_\text{VCO}=f_\text{min}$. So, if it's demanded a frequency of `{python} format(f_example_1, 3)`Hz, for example, it uses the following equation to determine the voltage of control, $V_c$

$$
\begin{split}
f_o&=f_\text{VCO}+K_\text{VCO}\cdot V_c\implies\\
V_c&=\frac{f_o-f_\text{VCO}}{K_\text{VCO}}
\end{split}
$$

V_c_example_1=(f_example_1-fvco)/Kvco
```

So the voltage control will be $V_c=$ `{python} format(V_c_example_1)`V.

@@

As it was commented before, the continuous-time analysis can be useful in this system. So the already known equations for calculation of the second-order loop filter. For the first capacitor, $C_1$, its value can be obtained as

$$
C_1=\frac{I_\text{CP}K_\text{VCO}}{2\pi\cdot N\omega_n^2}
$$

@@

C_1=5e-9
I_CP_ang=C_1*2*math.pi*N0*omega_n**2/Kvco_ang
I_CP=C_1*2*math.pi*N0*omega_n**2/Kvco

@@

The charge pump current, $I_\text{CP}$, is another unknown variable in this equation, so it's arbitraryly supposed a value of $C_1=$ `{python} format(C_1)`F, so the charge pump current should be $I_\text{CP}=$ `{python} format(I_CP_ang)`A. Then the resistor can be determined as

@@

R_1=2*zeta*math.sqrt((2*math.pi*N0)/(I_CP*Kvco*C_1))

Thus the resistor will be $R=$ `{python} format(R_1)`$\Omega$.

Using the common realtion, it's possible to find the second capacitor, $C_2$, for the second-order loop filter,

$$
C_2=\frac{C_1}{10}
$$

C_2=C_1/10 
```

Resulting in $C_2=$ `{python} format(C_2)`F.

>>>>

f0=3.7e9
f = 4.3e9
tunning_range = f-f0
step=1e6
fref = 40e6
I_CP=2*math.pi*100e-6
V_sup=3
f_3dB=150e3
omega_3dB=f_3dB*2*math.pi

Kvco=tunning_range/V_sup
Kvco_ang=Kvco*2*math.pi

f_30M=30e6
f_300M=300e6

Ktheta=I_CP/(2*math.pi)

fvco=4e9
N = fvco/fref
```

# Introdução
A `{python} format(f0)`Hz–`{python} format(f)`Hz synthesizer with a step size of `{python} format(step)`Hz is required. A `{python} format(fref)`Hz crystal oscillator, a charge pump with `{python} format(I_CP)`A output current, and a VCO (operating from a `{python} format(V_sup)`V supply) are available. Design a fractional-N synthesizer with a loop bandwidth of `{python} format(f_3dB)`Hz using these components. Estimate the settling time of the loop for a `{python} format(f_30M)`Hz and `{python} format(f_300M)`Hz frequency step. Simulate and compare.

@@

For the VCO, if the system is fed with a `{python} format(V_sup)`V supply and the tuning range is `{python} format(tunning_range)`Hz, so the VCO constant can be $K_\text{VCO}=$ `{python} format(Kvco)`Hz/V. And initially the nominal frequency is arbitrary selected as $f_\text{VCO}=$ `{python} format(fvco)`Hz. With that nominal frequency, the division ratio will be $N=$ `{python} N`.

## Charge pump

In addition, since the charge pump current is known, the charge pump constant can be determined as $K_\theta=$ `{python} format(Ktheta)`A/rad.

@@

@@

fref_f3dB=fref/f_3dB
```

As the ration of reference frequency and bandwidth is $f_\text{ref}/f_{3\text{dB}}=$ `{python} round(fref_f3dB, 0)`, i.e. $f_\text{ref}\gg f_{3\text{dB}}$, the continuous-time analysis can be used.

This section is responsible to present the design of the loop filter. First of all, it's needed to determine the value of natural frequency,

$$
\omega_n=\frac{\omega_{3\text{dB}}}{\sqrt{1+2\zeta^2+\sqrt{4\zeta^4+4\zeta^2+2}}}
$$

zeta = 0.707

omega_n=omega_3dB/math.sqrt(1+2*zeta**2+math.sqrt(4*zeta**4+4*zeta**2+2))
f_n=omega_n/(2*math.pi)
```

Choosing a damping constant, $\zeta=$ `{python} zeta`, $f_n=$ `{python} format(f_n)`Hz.

The first capacitor, $C_1$, can be calculated as

$$
C_1=\frac{IK_\text{VCO}}{2\pi\cdot N\omega_n^2}
$$

C_1=I_CP*Kvco_ang/(2*math.pi*N*omega_n**2)

@@

So this capacitor has a value of, $C_1=$ `{python} format(C_1)`F.

The first resistor (and the single one) used in the second-order loop filter will can be obtained as follows,

@@

R_1=2*zeta*math.sqrt((2*math.pi*N)/(I_CP*Kvco_ang*C_1))
C_2=C_1/10
```

The value of that first resistor is $R_1=$ `{python} format(R_1)`$\Omega$. And the second capacitor, $C_2$, is obtained as one-tenth of C_1, getting the value of $C_2=$ `{python} format(C_2)`F.

## Testing

t_m=1/(omega_n*math.sqrt(1-zeta**2))*math.atan(math.sqrt(1-zeta**2)/zeta)
```

theta_e_m_n = math.sin(omega_n*math.sqrt(1-zeta**2)*t_m)/(math.sqrt(1-zeta**2))*math.e**(-zeta*omega_n*t_m)
theta_e_m = 2 * math.pi
```

step_m_ang=theta_e_m*omega_n/theta_e_m_n
step_m_lin = step_m_ang/2/math.pi
```

The maximum input frequency step that can be tolerated for a system with these parameters is $\Delta\omega_m=$ `{python} format_with_pi_latex(step_m_ang)`rad/s, or `{python} format(step_m_lin)`Hz.

f_i_30M = f_30M/N
f_i_30M_ang = f_i_30M*2*math.pi
f_i_300M = f_300M/N
f_i_300M_ang = f_i_300M*2*math.pi
```

omega_n__t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta)*(-1)
t_s=omega_n__t_s/omega_n
```

Now, a step in output frequency of `{python} format(f_30M)`Hz and `{python} format(f_300M)`Hz corresponds to a step in the reference frequency (at the input) of `{python} format(f_i_30M)`Hz and `{python} format(f_i_300M)`Hz, respectively. The first frequency step must be linear once it is less than the maximum frequency step, and it must be settle in a time that can be calculated using the relationship obtained for $\omega_nt_s$, so the value of settling time must be $t_s=$ `{python} format(t_s)`s.

For the step of `{python} format(f_i_300M)`Hz, in the other hand, the step will involve cycle slipping. For this nonlinear case, it's needed to estimate the acquisition time. It can be done as follows

$$
T_s=\frac{\Delta\omega}{\pi\omega_n^2}
$$

Ts = f_i_300M_ang/(math.pi*omega_n**2)
Total_ts=Ts+t_s
```

Using the above equation, the acquisition time must be $T_s=$ `{python} format(Ts)`s, plus an aditional `{python} format(t_s)`s for phase lock. Therefore, the total time to phase lock, due to an step frequency of `{python} format(f_i_300M)`Hz is `{python} format(Total_ts)`s.
